
## Libraries
Libraries needed in the analyses
```{r, message=FALSE}
library(maps)
library(mapdata)
library(tidyverse)
library(gridExtra)
```

## Map of Europe and values for AR
The map of Europe is from the `mapdata` package. The antibiotic consumption data and the *E. coli* resistance rates (combined resistance: third-generation cephalosporin, fluoroquinolones and aminoglycoside) for year 2015 are from the [European Centre for Disease Prevention and Control.](http://ecdc.europa.eu)
```{r, message=FALSE}
EU_map <- map_data("world", region=c("Austria","Andorra","Belgium","Bulgaria","Croatia","Cyprus",
                   "Czech Rep.","Denmark","Estonia","Finland","France",
                   "Germany","Greece","Switzerland","Hungary","Italy", "Ireland", "Latvia",
                   "Liechtenstein", "Lithuania","Luxembourg","Malta","Netherlands", 
                   "Norway","Poland", "Portugal","Romania","Slovakia", "Slovenia",
                   "Spain", "Sweden", "UK"))

AB_data <- read_delim ("data/AB_data.txt", delim = "\t", col_names = TRUE)
Res_data <- read_csv("data/ECDC_surveillance_data_Antimicrobial_resistance.csv")
AR_rawdata <- read_delim("data/Stare_data.txt", delim=" ", col_names = TRUE)
AR_rawdata$region <- str_replace(AR_rawdata$id_plant, pattern="[:digit:]", replace ="")

AR_inf <- AR_rawdata %>% 
  filter(treatmentClassification=="untreated") %>% 
  group_by(region) %>% 
  summarize(inf_mean = mean(SUM)) %>% 
  ungroup()

AR_eff <- AR_rawdata %>% 
  filter(effluent==TRUE) %>% 
  group_by(region) %>% 
  summarize(eff_mean = mean(SUM)) %>% 
  ungroup()

AR_data <- full_join(AR_inf,AR_eff, by="region")

AR_data[AR_data$region=="PT",]$region <- "Portugal"
AR_data[AR_data$region=="CYP",]$region <- "Cyprus"
AR_data[AR_data$region=="FI",]$region <- "Finland"
AR_data[AR_data$region=="DE",]$region <- "Germany"
AR_data[AR_data$region=="NO",]$region <- "Norway"
AR_data[AR_data$region=="ES",]$region <- "Spain"
AR_data[AR_data$region=="IL",]$region <- "Ireland"

EU_map <- full_join(EU_map, AR_data, by="region")
EU_map <- full_join(EU_map, AB_data, by="region")
EU_map <- full_join(EU_map, Res_data, by="region")

EU_map <- EU_map %>% 
  mutate(scaled_DDD = as.vector(scale(DDD)), 
         scaled_res = as.vector(scale(NumValue)), 
         scaled_inf = as.vector(scale(inf_mean)), 
         scaled_eff = as.vector(scale(eff_mean)))
```
## Figures
```{r, fig.width=10, fig.height=4}
p1 <- ggplot() + 
  geom_polygon(data = EU_map, aes(x=long, y = lat, group = group, fill=scaled_DDD), color="grey40") + 
  scale_y_continuous(limits = c(34,71)) + 
  theme(panel.grid = element_blank(), panel.background = element_blank(),
        plot.title = element_text(size=10, face = "bold"), legend.position = "none") +
  labs(x="Longitude", y="Latitude", title = "Antibiotic use in Europe", caption = "source: ecdc.europa.eu") + 
  scale_fill_gradient(name="scale", high="red", low="#FFFF99", na.value = "grey80")

p2 <- ggplot() + 
  geom_polygon(data = EU_map, aes(x=long, y = lat, group = group, fill=scaled_res), color="grey40") + 
  scale_y_continuous(limits = c(34,71)) + 
  theme(panel.grid = element_blank(), panel.background = element_blank(), 
        plot.title = element_text(size=10, face = "bold"), legend.position = "none") +
  labs(x="Longitude", y="Latitude", title = "Clinical resistance (E. coli)", caption = "source: ecdc.europa.eu") + 
  scale_fill_gradient(name="E. coli\nresistant isolates", high="red", low="#FFFF99", na.value = "grey80")

p3 <- ggplot() + 
  geom_polygon(data = EU_map, aes(x=long, y = lat, group = group, fill=scaled_inf), color="grey40") + 
  scale_y_continuous(limits = c(34,71)) + 
  theme(panel.grid = element_blank(), panel.background = element_blank(), 
        plot.title = element_text(size=10, face = "bold"), legend.position = "none") +
  labs(x="Longitude", y="Latitude", title = "ARGs in raw sewage", caption = "source: this study") + 
  scale_fill_gradient(name="Influent\nAR prevalence", high="red", low="#FFFF99", na.value = "grey80")

grid.arrange(p1, p2, p3, ncol=3)
```

**Figure 1.** Antibiotic consumption in European countries as DDD per 1 000 inhabitants (**left**). Proportion of clinical antibiotic resistant (AR) _E. coli_ isolates to third-generation cephalosporin, fluoroquinolones and aminoglycoside (**center**). Antibiotic resistance genes (ARGs) detected in raw sewage from countries included in this study (**right**). Data on antibiotic onsumption and resistant _E. coli_ is for year 2015 and from ECDC. Antibiotic resistance gene data from this study. All data is scaled. Yellow color designates low values and red color designates high values.